---
title: "Exploratory Data Analysis for Calibration"
output: 
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: true
editor_options: 
  chunk_output_type: console
params:
  variable: 'SWCF_LWCF_PRECT_PSL_Z500_U200_U850_TREFHT_U_RELHUM_T_RESTOM_0.7'
  resolution: '180x360'
  nyears: '10yr'
  season: 'DJF'
  type: 'lat_lon'
  subtract: 'raw'
  n_components: 16
---

\pagebreak

# Preliminaries

Load in packages and information on what document we are making.
```{r setup, message = FALSE}
library(knitr)
knitr::opts_chunk$set(cache=F, dev = 'png', warning = FALSE, echo = FALSE,
                      message = FALSE, cache.lazy = FALSE)
library(ncdf4)
library(fields)
print(unlist(params))
variable <- params[['variable']]
variables <- strsplit(variable, '_')[[1]]
if ('RESTOM' %in% variables) {
  variables <- variables[-length(variables)]
}
variable_types <- c('SWCF' = 'lat_lon', 
                    'LWCF' = 'lat_lon',
                    'PRECT' = 'lat_lon',
                    'PSL' = 'lat_lon',
                    'Z500' = 'lat_lon',
                    'U200' = 'lat_lon',
                    'U850' = 'lat_lon',
                    'TREFHT' = 'lat_lon',
                    'U' = 'lat_plev',
                    'T' = 'lat_plev',
                    'RELHUM' = 'lat_plev',
                    'RESTOM' = 'global')
unique_variable_types  <-  unique(variable_types[variables])
variables_plev <- variables[variable_types[variables] == 'lat_plev']
variables_lon <- variables[variable_types[variables] == 'lat_lon']
variables_global <- variables[variable_types[variables] == 'global']
n_plev_vars  <- length(variables_plev)
n_lon_vars  <- length(variables_lon)

res_vals_lon <- as.numeric(strsplit(params[['resolution']], 'x')[[1]])
res_vals_plev <- c(24, 37)
nvar <- length(variables)
if (nvar == 3) {
  plot_dim_im = list(mfrow=c(2,2), mar= 4*c(0.5, 0.5, 0.5, 0.5))
  plot_dim = list(mfrow=c(2,2), mar= 8*c(0.5, 0.5, 0.5, 0.5))
} else if (nvar == 1) {
  plot_dim_im = list(mfrow=c(1,1))
  plot_dim = list(mfrow=c(1,1))
} else  if (nvar == 8) {
  plot_dim_im = list(mfrow=c(3,3), mar= 4*c(0.5, 0.5, 0.5, 0.5))
  plot_dim = list(mfrow=c(3,3), mar= 4*c(0.5, 0.5, 0.5, 0.5))
}
```

## Load model outputs

```{r}
# get simulation runs created by Kenny's code
if (params[['season']] == 'ALL') {
  eg <- expand.grid(variables[variables != 'RESTOM'], c('DJF', 'MAM', 'JJA', 'SON'),
                    stringsAsFactors = F)
  if ('RESTOM' %in% variables) {
    eg <- rbind(eg, c('RESTOM', 'ANN'))
  }
  variables_vec <- as.character(eg[,1])
  seasons_vec <- eg[,2]
  variables <- sprintf('%s: %s', eg[,1], eg[,2])
  n_plev_vars  <- sum(variable_types[variables_vec] == 'lat_plev')
  n_lon_vars  <- sum(variable_types[variables_vec] == 'lat_lon')
  nvar <- n_plev_vars + n_lon_vars + 
    'RESTOM' %in% variables
} else {
  seasons_vec <- rep(params[['season']], nvar)
  variables_vec <- variables
}
misc_runs <- c('0522', '8fields_ALL_24x48', '8fields_ANN_24x48', 
               '11fields_ALL_24x48')

data_frame_list <- list()
sim_tensor <- array(dim = c(180, 360, 250, 8, 4))
#sur_tensor <- array(dim = c(180, 360, 250, 8, 4))
data_frame_pcs_list <- list()

for (j in 1:length(unique_variable_types)) {
    surr_file <- nc_open(paste0('surrogate_models/pred_',params[['season']], '_',  params[['nyears']], '_obs_full_', params[['resolution']], '_', params[['n_components']], '_',  params[['variable']], '_', params[['subtract']], '.nc'))
    
    # load principal components
    pc_file <- nc_open(paste0('surrogate_models/pcs_',params[['season']], '_',  params[['nyears']], '_obs_full_',
                          params[['resolution']], '_', params[['n_components']], '_',  params[['variable']], '_', params[['subtract']], '.nc'))
    pc_scores_model <- ncvar_get(pc_file, 'pc_scores_model')
    pc_scores_surr <- ncvar_get(pc_file, 'pc_scores_surr')
    pc_scores_obs <- ncvar_get(pc_file, 'pc_scores_obs')
    pc_vals <- ncvar_get(pc_file, 'pc_vals')
    pc_prop_var <- ncvar_get(pc_file, 'prop_var')
    ev <- ncvar_get(pc_file, 'ev')
    sing_vals <- ncvar_get(pc_file, 'sing_vals')
    nc_close(pc_file)
    
    variables_use <- variables_vec[variable_types[variables_vec] == unique_variable_types[j]]
    seasons_use <- seasons_vec[variable_types[variables_vec] == unique_variable_types[j]]
    if (unique_variable_types[j] == 'global') {
          vals_use <- ncvar_get(surr_file, 'values')
    data_frame_list[[length(data_frame_list) + 1]] <- data.frame('season' = NA, lon = NA, lat = NA, plev  = NA, type  = 'surr', 
               sim_run = as.character(1:250), variable = 'RESTOM', res = NA,
               value = vals_use[nrow(vals_use),]) 
    nc_close(surr_file)
      sim_file <- nc_open(paste0('data/lat_lon_', params[['nyears']], 
                                 '_180x360_', seasons_use, '.nc'))
      FLNT <- ncvar_get(sim_file, 'FLNT')
      FSNT <- ncvar_get(sim_file, 'FSNT')
      area <- ncvar_get(sim_file, 'area')
      area_norm <-  area/sum(area) * 250
      RESTOM <- apply(area_norm * (FSNT - FLNT),3, sum)
      data_frame_list[[length(data_frame_list) + 1]] <- data.frame('season' = NA, lon = NA, lat = NA, plev  = NA, type  = 'sim', 
                                                                 sim_run = as.character(1:250), variable = 'RESTOM', res = NA,
                                                                 value = RESTOM) 
      ctrl_file <- nc_open(paste0('data/lat_lon_', params[['nyears']], '_180x360_', 
                           seasons_use, '_ctrl.nc'))
      FLNT <- ncvar_get(ctrl_file, 'FLNT')
      FSNT <- ncvar_get(ctrl_file, 'FSNT')
      area <- ncvar_get(ctrl_file, 'area')
      nc_close(ctrl_file)
      area_norm <-  area/sum(area) 
      RESTOM_ctrl <- sum(area_norm * (FSNT - FLNT))
      data_frame_list[[length(data_frame_list) + 1]] <- data.frame('season' = NA, lon = NA, lat = NA, plev  = NA, type  = 'ctrl', 
                       sim_run = NA, variable = 'RESTOM', res = NA, value = RESTOM_ctrl)
      
      RESTOM_misc <- rep(0, length(misc_runs))
      for (r in 1:length(misc_runs)) {
        misc_file <- nc_open(paste0('data/lat_lon_10yr_180x360_ANN_miscsim_',
                                  misc_runs[r], '.nc'))
        FLNT <- ncvar_get(misc_file, 'FLNT')
        FSNT <- ncvar_get(misc_file, 'FSNT')
        area <- ncvar_get(misc_file, 'area')
        area_norm <-  area/sum(area) 
        RESTOM_misc[r] <- sum(area_norm * (FSNT - FLNT))
        nc_close(misc_file)
      }
      names(RESTOM_misc) <- misc_runs
      data_frame_list[[length(data_frame_list) + 1]] <- data.frame('season' = NA, lon = NA, lat = NA, plev  = NA, type  = misc_runs,
                       sim_run = NA, variable = 'RESTOM', res = NA, value = RESTOM_misc)
      # PCs
      data_frame_pcs_list[[length(data_frame_pcs_list) + 1]] <- 
         data.frame('season' = 'ANN', lon = NA, lat = NA, plev  = NA, pc = 1:(dim(pc_vals)[2]),
                    variable = 'RESTOM', res = q,
                   value = as.vector(pc_vals[nrow(pc_vals),]) )
      next
    }
    for (q in c('24x48', '180x360')) {
      if (unique_variable_types[j] == 'lat_plev' & q == '180x360') {
        next
      } #else if (unique_variable_types[j] == 'lat_plev') {
      #  q = '24x37'
      #}
      for (i in 1:length(variables_use)) {
      sim_file <- nc_open(paste0('data/', unique_variable_types[j], '_', params[['nyears']], '_',
                           q, '_', 
                           seasons_use[i], '.nc'))
      obs_file <- nc_open(paste0('data/', unique_variable_types[j], '_',  q, '_', seasons_use[i], '_obs.nc'))
      ctrl_file <- nc_open(paste0('data/', unique_variable_types[j], '_', params[['nyears']], '_', 
                            q, '_', seasons_use[i], '_ctrl.nc'))
      if (unique_variable_types[j] == 'lat_lon') {
        lon <- sim_file[['dim']][['lon']][['vals']]
        lat <- sim_file[['dim']][['lat']][['vals']]
        lat_rep <- lat_surr_rep <- rep(lat, each = length(lon))
        plev <- plev_surr <- NA
        sim_run <- rep(1:250, each = length(lon) * length(lat)) 
        indexes_use  <- list(1:length(lon), 1:length(lat))
        indexes_variable_all <- (1 + (i-1)*prod(res_vals_lon)):
    (i*prod(res_vals_lon))
        if (q == params[['resolution']]) {
          surr_pred <- ncvar_get(surr_file, 'values')[indexes_variable_all,]
        }
        pcs_use <- pc_vals[indexes_variable_all,]
      } else {
        mask <- ncvar_get(surr_file, 'mask')
        plev <- sim_file[['dim']][['plev']][['vals']]
        lon <- NA
        lat <- sim_file[['dim']][['lat']][['vals']]
        lat_surr_rep <- rep(lat, each =  length(plev))
        plev_surr <-  plev
        lat_rep <- lat
        plev <- rep(plev, each = length(lat_rep))
        sim_run <- rep(1:250, each = length(plev_surr) * length(lat)) 
        indexes_variable_all <- (1 + n_lon_vars*(res_vals_lon[2] * res_vals_lon[1]) +
                                    sum(mask[,1:i]) - sum(mask[,i])):
    (n_lon_vars*(res_vals_lon[2] * res_vals_lon[1]) + sum(mask[,1:i]))
        surr_pred_prelim <- ncvar_get(surr_file, 'values')[indexes_variable_all,]
        surr_pred <- matrix(NA, nrow = prod(res_vals_plev), ncol = 250)
        surr_pred[as.logical(mask[,i]), ] <- surr_pred_prelim
        
        pcs_prelim <- pc_vals[indexes_variable_all,]
        pcs_use <- matrix(NA, nrow = prod(res_vals_plev), ncol = 16)
        pcs_use[as.logical(mask[,i]), ] <- pcs_prelim
      }
      # simulations
      if (q != '180x360') {
        data_frame_list[[length(data_frame_list) + 1]] <- 
       data.frame('season' = seasons_use[i], lon, lat = lat_rep, plev, type  = 'sim', sim_run = as.character(sim_run), variable = variables_use[i], 
                  res = q,
                   value = as.vector(ncvar_get(sim_file, variables_use[i])))
      } else {
        sim_tensor[,, ,which(variables_lon == variables_use[i]),which(levels(seasons_use) == seasons_use[i])] <- ncvar_get(sim_file, variables_use[i])
      }
      
      # surrogate
      if (params[['resolution']] == q) {
              data_frame_list[[length(data_frame_list) + 1]] <- 
        data.frame('season' = seasons_use[i], lon = (lon), lat = (lat_surr_rep), plev  = (plev_surr), type  = 'surr', 
                   sim_run = as.character(sim_run), variable = variables_use[i], res = q,
                   value = as.vector(surr_pred)) 
      }

      # obs
      data_frame_list[[length(data_frame_list) + 1]] <- 
        data.frame('season' = seasons_use[i], lon = (lon), lat = (lat_rep), 
                   plev = (plev), type  = 'obs',  sim_run =NA, variable = variables_use[i], res = q,
                   value =  as.vector(ncvar_get(obs_file, variables_use[i])))
      
      # ctrl
      data_frame_list[[length(data_frame_list) + 1]] <- 
        data.frame('season' = seasons_use[i],lon = (lon), lat = (lat_rep), 
                   plev = (plev), type = 'ctrl', sim_run =NA, variable = variables_use[i], res = q,
                   value =  as.vector(ncvar_get(ctrl_file, variables_use[i])))
      # misc
      for (k in 1:length(misc_runs)) {
        misc_file <- nc_open(paste0('data/', unique_variable_types[j], '_', params[['nyears']], '_', 
                        q, '_',seasons_use[i], '_miscsim_', misc_runs[k], '.nc'))
        data_frame_list[[length(data_frame_list) + 1]] <- 
          data.frame('season' = seasons_use[i], lon = (lon), lat = (lat_rep), 
                   plev = (plev), type = misc_runs[k], sim_run =NA, variable = variables_use[i], res = q,
                   value =as.vector(ncvar_get(misc_file, as.character(variables_use[i]))))
        nc_close(misc_file)
      }
      lhs <- base::t(ncvar_get(sim_file, 'lhs'))
      input_variables <- ncvar_get(sim_file, 'x')
      nc_close(sim_file)
      nc_close(obs_file)
      nc_close(ctrl_file)
      
                #pc values
      if (q == params[['resolution']])
    data_frame_pcs_list[[length(data_frame_pcs_list) + 1]] <- 
         data.frame('season' = seasons_use[i], lon, lat = lat_surr_rep, plev  = plev_surr, pc = rep(1:(dim(pc_vals)[2]), each = length(lat_surr_rep)),
                    variable = variables_use[i], res = q,
                   value = as.vector(pcs_use) )
    }
    }
    
  }
df <- dplyr::bind_rows(data_frame_list)

colnames(lhs) <- input_variables
```

## Load area file for weighting
```{r}
area_file <- nc_open(paste0('surrogate_models/area_24x48.nc'))
area <- ncvar_get(area_file, 'values')
sim_file_test <- nc_open(paste0('data/lat_lon_', params[['nyears']], '_24x48_', 
                           'ANN', '.nc'))

area_lat <- sim_file_test$dim$lat$vals
area_lon <- sim_file_test$dim$lon$vals
area_norm <- area/mean(area)
nc_close(area_file)
nc_close(sim_file_test)

area_df <- data.frame(area_wt = as.vector(area_norm),
                      lat = rep(area_lat, each=  length(area_lon)), 
                      lon = area_lon, 
                      res = '24x48')
area_file <- nc_open(paste0('surrogate_models/area_180x360.nc'))
area <- ncvar_get(area_file, 'values')
sim_file_test <- nc_open(paste0('data/lat_lon_', params[['nyears']], '_180x360_', 
                           'ANN', '.nc'))

area_lat <- sim_file_test$dim$lat$vals
area_lon <- sim_file_test$dim$lon$vals
area_norm <- area/mean(area)
nc_close(area_file)
nc_close(sim_file_test)
area_df2 <- data.frame(area_wt = as.vector(area_norm),
                      lat = rep(area_lat, each=  length(area_lon)), 
                      lon = area_lon, 
                      res = '180x360')
area_df <- rbind(area_df, area_df2)

```


## Subtract the ensemble mean if necessary

If we ran the surrogate after subtracting the ensemble mean, adjust the simulation output and the observations accordingly. 

```{r}
library(dplyr)
sim_avg <- df  %>%
  filter(type == 'sim') %>%
  group_by(lon, lat, plev, res) %>%
  summarize(mv = mean(value))
df <- left_join(df, sim_avg)
if (params[['subtract']] == 'ensmean') {
  df[['value']][df[['type']] != 'surr'] <- df[['value']][df[['type']] != 'surr'] -
    df[['mv']][df[['type']] != 'surr']
}
```


# A few initial plots just to get the feel of things

## The observational fields
```{r, dpi = 144}
library(purrr)
library(cowplot)
library(ggplot2)
df %>% 
  filter(type == 'obs', !is.na(lon), res == '24x48') %>%
  group_split(variable) %>% 
  map(
    ~ggplot(., aes(lon, lat, fill = value)) + 
      geom_raster() +
      scale_fill_gradientn(colors = rainbow(10)) + 
      facet_wrap(~ variable + season, labeller = function(x) label_value(x, multi_line = FALSE))+
                   theme(legend.title = element_blank(),
                         axis.title = element_blank(),
                         axis.ticks = element_blank(),
                         axis.text = element_blank(),
                         legend.key.size = unit(.5, "lines"),
                         legend.margin=margin(0,0,0,0),
                         legend.box.margin=margin(-10,-10,-10,-10),
                         legend.text = element_text(size = 5))
  ) %>% 
  plot_grid(plotlist = ., align = 'hv')


df %>% 
  filter(type == 'obs', !is.na(lon), res == '180x360') %>%
  group_split(variable) %>% 
  map(
    ~ggplot(., aes(lon, lat, fill = value)) + 
      geom_raster() +
      scale_fill_gradientn(colors = rainbow(10)) + 
      facet_wrap(~ variable + season, labeller = function(x) label_value(x, multi_line = FALSE))+
                   theme(legend.title = element_blank(),
                         axis.title = element_blank(),
                         axis.ticks = element_blank(),
                         axis.text = element_blank(),
                         legend.key.size = unit(.5, "lines"),
                         legend.margin=margin(0,0,0,0),
                         legend.box.margin=margin(-10,-10,-10,-10),
                         legend.text = element_text(size = 5))
  ) %>% 
  plot_grid(plotlist = ., align = 'hv')

plevels <- unique(df$plev)[!is.na(unique(df$plev))]
plevels_df <- data.frame(plev = plevels, 
                      plev_lower = lead(plevels))
plevels_df$plev_lower[is.na(plevels_df$plev_lower)] <- 0

lats <- unique(df$lat)
lat_lag <- lats[2] - lats[1]
lats_df <- data.frame(lat = lats,
                      lat_higher = lats + lat_lag/2, 
                      lat_lower = lats - lat_lag/2)

df %>% 
  filter(type == 'obs', !is.na(plev)) %>%
  left_join(plevels_df) %>%
  left_join(lats_df) %>% 
  group_split(variable) %>% 
  map(
    ~ggplot(., aes(ymin = plev_lower, ymax = plev, xmin = lat_lower, xmax = lat_higher, color = value, fill = value)) + 
      geom_rect() +
      scale_color_gradientn(colors = rainbow(10)) + 
      scale_fill_gradientn(colors = rainbow(10)) + 
      scale_y_continuous(trans = 'reverse') + 
      facet_wrap(~ variable + season, labeller = function(x) label_value(x, multi_line = FALSE))+
                   theme(legend.title = element_blank(),
                         axis.title = element_blank(),
                         axis.ticks = element_blank())
  ) %>% 
  plot_grid(plotlist = ., align = 'hv')
```

## One simulation run

```{r, dpi = 144}
df %>% 
  filter(type == 'sim', !is.na(lon), sim_run == 20, res == '24x48') %>%
  group_split(variable) %>% 
  map(
    ~ggplot(., aes(lon, lat, fill = value)) + 
      geom_raster() +
      scale_fill_gradientn(colors = rainbow(10)) + 
      facet_wrap(~ variable + season, labeller = function(x) label_value(x, multi_line = FALSE))+
                   theme(legend.title = element_blank(),
                         axis.title = element_blank(),
                         axis.ticks = element_blank(),
                         axis.text = element_blank(),
                         legend.key.size = unit(.5, "lines"),
                         legend.margin=margin(0,0,0,0),
                         legend.box.margin=margin(-10,-10,-10,-10),
                         legend.text = element_text(size = 5))
  ) %>% 
  plot_grid(plotlist = ., align = 'hv')


df %>% 
  filter(type == 'sim', !is.na(plev), sim_run == 20) %>%
  left_join(plevels_df) %>%
  left_join(lats_df) %>% 
  group_split(variable) %>% 
  map(
    ~ggplot(., aes(ymin = plev_lower, ymax = plev, xmin = lat_lower, xmax = lat_higher, color = value, fill = value)) + 
      geom_rect() +
      scale_color_gradientn(colors = rainbow(10)) + 
      scale_fill_gradientn(colors = rainbow(10)) + 
      scale_y_continuous(trans = 'reverse') + 
      facet_wrap(~ variable + season, labeller = function(x) label_value(x, multi_line = FALSE))+
                   theme(legend.title = element_blank(),
                         axis.title = element_blank(),
                         axis.ticks = element_blank())
  ) %>% 
  plot_grid(plotlist = ., align = 'hv')
```

## The surrogate output

```{r, dpi = 144}
df %>% 
  filter(type == 'surr', !is.na(lon), sim_run == 20, res == '24x48') %>%
  group_split(variable) %>% 
  map(
    ~ggplot(., aes(lon, lat, fill = value)) + 
      geom_raster() +
      scale_fill_gradientn(colors = rainbow(10)) + 
      facet_wrap(~ variable + season, labeller = function(x) label_value(x, multi_line = FALSE))+
                   theme(legend.title = element_blank(),
                         axis.title = element_blank(),
                         axis.ticks = element_blank(),
                         axis.text = element_blank(),
                         legend.key.size = unit(.5, "lines"),
                         legend.margin=margin(0,0,0,0),
                         legend.box.margin=margin(-10,-10,-10,-10),
                         legend.text = element_text(size = 5))
  ) %>% 
  plot_grid(plotlist = ., align = 'hv')


df %>% 
  filter(type == 'surr', !is.na(plev), sim_run == 20) %>%
  left_join(plevels_df) %>%
  left_join(lats_df) %>% 
  group_split(variable) %>% 
  map(
    ~ggplot(., aes(ymin = plev_lower, ymax = plev, xmin = lat_lower, xmax = lat_higher, color = value, fill = value)) + 
      geom_rect() +
      scale_color_gradientn(colors = rainbow(10)) + 
      scale_fill_gradientn(colors = rainbow(10)) + 
      scale_y_continuous(trans = 'reverse') + 
      facet_wrap(~ variable + season, labeller = function(x) label_value(x, multi_line = FALSE))+
                   theme(legend.title = element_blank(),
                         axis.title = element_blank(),
                         axis.ticks = element_blank())
  ) %>% 
  plot_grid(plotlist = ., align = 'hv')
```

## Simulation run minus surrogate

```{r, dpi = 144}
df %>%
  filter(type == 'surr', !is.na(lon), sim_run == 20, res == '24x48') %>%
  left_join(filter(df, type == 'sim', !is.na(lon), sim_run == 20), by = c('season',  'lon', 'lat','plev', 'sim_run', 'variable')) %>%
    group_split(variable) %>% 
  map(
    ~ggplot(., aes(lon, lat, fill = value.y - value.x)) + 
      geom_raster() +
      scale_fill_gradientn(colors = rainbow(10)) + 
      facet_wrap(~ variable + season, labeller = function(x) label_value(x, multi_line = FALSE))+
                   theme(legend.title = element_blank(),
                         axis.title = element_blank(),
                         axis.ticks = element_blank(),
                         axis.text = element_blank(),
                         legend.key.size = unit(.5, "lines"),
                         legend.margin=margin(0,0,0,0),
                         legend.box.margin=margin(-10,-10,-10,-10),
                         legend.text = element_text(size = 5))
  ) %>% 
  plot_grid(plotlist = ., align = 'hv')

df %>%
  filter(type == 'surr', !is.na(plev), sim_run == 20) %>%
  left_join(filter(df, type == 'sim', !is.na(plev), sim_run == 20), by = c('season',  'lon', 'lat','plev', 'sim_run', 'variable')) %>%
  left_join(plevels_df) %>%
  left_join(lats_df) %>% 
  group_split(variable) %>% 
  map(
    ~ggplot(., aes(ymin = plev_lower, ymax = plev, xmin = lat_lower, xmax = lat_higher, color = value.y - value.x, fill = value.y - value.x)) + 
      geom_rect() +
      scale_color_gradientn(colors = rainbow(10)) + 
      scale_fill_gradientn(colors = rainbow(10)) + 
      scale_y_continuous(trans = 'reverse') + 
      facet_wrap(~ variable + season, labeller = function(x) label_value(x, multi_line = FALSE))+
                   theme(legend.title = element_blank(),
                         axis.title = element_blank(),
                         axis.ticks = element_blank())
  ) %>% 
  plot_grid(plotlist = ., align = 'hv')
```

```{r, dpi = 144}
library(rcolors)
WBGYR <- rcolors:: get_color('WhiteBlueGreenYellowRed', n = NULL)
df %>%
  filter(type %in% c('surr', 'sim'), !is.na(lon), sim_run == 20, res == '24x48', variable == 'PRECT', season == 'JJA') %>%
  left_join(data.frame(type = c('surr', 'sim'), 
                       type_label = factor(c('Surrogate prediction', 
                                             'E3SM output')))) %>%
  ggplot(aes(ifelse(lon > 180, lon - 360, lon), lat, fill = value)) + 
      geom_raster() +
  geom_polygon(data = map_data('world'),
               aes(x = long, y = lat, group = group),
               fill = NA, color = 'gray70', size = .1) + 
  coord_quickmap() + 
  scale_x_continuous(breaks = c(-180, -90, 0, 90, 180)) +
  scale_y_continuous(breaks = c(-90, -60, -30, 0, 30, 60, 90)) + 
      scale_fill_gradientn(colours = WBGYR) + 
      facet_wrap(~type_label, ncol = 2)+
  theme_bw() +
  theme( strip.background = element_blank(),
         strip.text.x = element_blank(),
         legend.position = 'bottom') + 
  labs(x = 'Longitude', y = 'Latitude',
       fill = 'PRECT\n(mm/day)')
ggsave(paste0('src/eda/paper_plots/', params[['variable']], '_prect_JJA.png'),
       width = 10, height = 4)
BrBG <- rcolors:: get_color('BrBG', n = NULL)
prelim_df <- df %>%
  filter(type %in% c('surr', 'sim'), !is.na(lon), sim_run == 20, res == '24x48', variable == 'PRECT', season == 'JJA') %>%
  tidyr::pivot_wider(names_from = type, values_from = value) %>%
  mutate(diff  = sim - surr)
ggplot(data = prelim_df, aes(ifelse(lon > 180, lon - 360, lon), lat, fill = diff)) + 
      geom_raster() +
  geom_polygon(data = map_data('world'),
               aes(x = long, y = lat, group = group),
               fill = NA, color = 'gray70', size = .1) + 
  coord_quickmap() + 
  scale_x_continuous(breaks = c(-180, -90, 0, 90, 180)) +
  scale_y_continuous(breaks = c(-90, -60, -30, 0, 30, 60, 90)) + 
  scale_fill_gradientn(colors = BrBG, limits = c(-max(abs(prelim_df$diff)), max(abs(prelim_df$diff))) ) + 
  theme_bw() +
  labs(x = 'Longitude', y = 'Latitude',
       fill = 'PRECT\n(mm/day)')
ggsave(paste0('src/eda/paper_plots/', params[['variable']], '_prect_JJA_diff.png'),
       width = 5.5, height = 2.5)



df %>%
  filter(type == 'surr', !is.na(plev), sim_run == 20) %>%
  left_join(filter(df, type == 'sim', !is.na(plev), sim_run == 20), by = c('season',  'lon', 'lat','plev', 'sim_run', 'variable')) %>%
  left_join(plevels_df) %>%
  left_join(lats_df) %>% 
  group_split(variable) %>% 
  map(
    ~ggplot(., aes(ymin = plev_lower, ymax = plev, xmin = lat_lower, xmax = lat_higher, color = value.y - value.x, fill = value.y - value.x)) + 
      geom_rect() +
      scale_color_gradientn(colors = rainbow(10)) + 
      scale_fill_gradientn(colors = rainbow(10)) + 
      scale_y_continuous(trans = 'reverse') + 
      facet_wrap(~ variable + season, labeller = function(x) label_value(x, multi_line = FALSE))+
                   theme(legend.title = element_blank(),
                         axis.title = element_blank(),
                         axis.ticks = element_blank())
  ) %>% 
  plot_grid(plotlist = ., align = 'hv')
```


## A plot at one location of the variable over the 250 runs

```{r, dpi = 144}
df_temp <- df  %>% filter(!is.na(lon), res == '24x48') %>% filter(lon == lon[20], lat == lat[20]) %>%
         mutate(sim_run = ifelse(is.na(as.numeric(sim_run)), 0, as.numeric(sim_run)))
  ggplot() + 
      geom_point(data= filter(df_temp, type %in% c('sim', 'surr')), aes(sim_run, y = value, color = type), size = .5) +
      geom_point(data = filter(df_temp, !(type %in% c('sim', 'surr'))), aes(sim_run, y = value, color = type), size = 2) +
      facet_grid(variable~ season, labeller = function(x) label_value(x, multi_line = FALSE), scales = 'free_y') +  
  labs(size = 'Size')
  
  
  df_temp <- df  %>% filter(!is.na(plev)) %>% filter(plev == plev[20], lat == lat[20]) %>%
         mutate(sim_run = ifelse(is.na(as.numeric(sim_run)), 0, as.numeric(sim_run)))
  ggplot() + 
      geom_point(data= filter(df_temp, type %in% c('sim', 'surr')), aes(sim_run, y = value, color = type), size = .5) +
      geom_point(data = filter(df_temp, !(type %in% c('sim', 'surr'))), aes(sim_run, y = value, color = type), size = 2) +
      facet_grid(variable~ season, labeller = function(x) label_value(x, multi_line = FALSE), scales = 'free_y') +  
  labs(size = 'Size')

```

# Compute R-squared at each location


Instead of averaging over locations, what if we average over simulation runs and see how well we capture between-simulation variability at each location. 

Formally, let $Y_i^{(\tau)}$ be the $i$-th location for the $\tau$-th simulation run. We take \begin{align*}
R^2(i) &= 1 - \frac{\frac{1}{250}\sum_{\tau=1}^{250}\left(Y_i^{(\tau)} - \hat{Y}_i^{(\tau)}\right)^2}{\frac{1}{250}\sum_{\tau=1}^{250}\left(Y_i^{(\tau)} - \overline{Y}_i\right)^2}
\end{align*}where $\overline{Y}_i = \frac{1}{250}\sum_{\tau=1}^{250} Y_i^{(\tau)}$.

```{r, dpi = 144}
surr_minus_sim <- df %>%
  dplyr::filter(res == '24x48', type %in% c('surr', 'sim')) %>%
  tidyr::pivot_wider(names_from = 'type') %>%
  mutate(surr_minus_sim = surr - sim) %>%
  left_join(area_df)
MSE <- surr_minus_sim %>%
  dplyr::filter(!is.na(lon)) %>%
  group_by(sim_run, variable,  season) %>%
  summarize(surr_error = mean(area_wt * (surr_minus_sim)^2),
            mean_error = mean(area_wt * (sim - mean(sim))^2))
MSE_loc <- surr_minus_sim %>%
  group_by(lat, lon, plev, variable,season) %>%
  summarize(surr_error = mean(surr_minus_sim^2),
            mean_error = mean((sim - mean(sim))^2),
            r2 = 1 - surr_error/mean_error)
library(cetcolor)
cet_perp_unif <- cetcolor::cet_pal('r2')
MSE_loc %>%
  filter(!is.na(lon)) %>%
  ggplot(aes(x= ifelse(lon > 180, lon - 360, lon), y  = lat, fill = r2)) +
  geom_raster() +
  geom_polygon(data = map_data('world'),
               aes(x = long, y = lat, group = group),
               fill = NA, color = 'gray10', size = .07) + 
  #scale_fill_viridis_c(limits = c(0,1)) + 
  scale_fill_gradientn(limits = c(0,1), colors = cet_perp_unif) + 
  coord_quickmap() + 
  facet_grid(variable~season) + 
  theme_bw() + 
  labs(x = 'Longitude', y = 'Latitude', fill = 'R-squared') +
  scale_x_continuous(breaks = c(-180, -90, 0, 90)) +
  scale_y_continuous(breaks = c(-90, -45, 0, 45, 90)) + 
  theme(axis.text = element_text(size = 8),
        legend.position = 'bottom')
ggsave(paste0('src/eda/paper_plots/', params[['variable']], '_r2_lat_lon.png'),
       width = 6, height = 7)

MSE_loc %>%
  filter(!is.na(lon), variable %in% c('PRECT', 'LWCF', 'U200')) %>%
  ggplot(aes(x= ifelse(lon > 180, lon - 360, lon), y  = lat, fill = r2)) +
  geom_raster() +
  geom_polygon(data = map_data('world'),
               aes(x = long, y = lat, group = group),
               fill = NA, color = 'gray10', size = .07) + 
  #scale_fill_viridis_c(limits = c(0,1)) + 
  scale_fill_gradientn(limits = c(0,1), colors = cet_perp_unif) + 
  coord_quickmap() + 
  facet_grid(variable~season) + 
  theme_bw() + 
  labs(x = 'Longitude', y = 'Latitude', fill = 'R-squared') +
  scale_x_continuous(breaks = c(-180, -90, 0, 90)) +
  scale_y_continuous(breaks = c(-90, -45, 0, 45, 90)) + 
  theme(axis.text = element_text(size = 8),
        legend.position = 'bottom')
ggsave(paste0('src/eda/paper_plots/', params[['variable']], '_r2_lat_lon_reduced.png'),
       width = 6, height = 3.75)

MSE_loc %>%
  filter(!is.na(plev)) %>%
  left_join(plevels_df) %>%
  left_join(lats_df) %>%
ggplot(., aes(ymin = plev_lower/100, ymax = plev/100, xmin = lat_lower, xmax = lat_higher, color = r2, fill = r2)) + 
      geom_rect() +
    scale_fill_gradientn(limits = c(0,1), colors = cet_perp_unif) + 
    scale_color_gradientn(limits = c(0,1), colors = cet_perp_unif) + 
  #scale_fill_viridis_c(limits = c(0,1)) + 
  #scale_color_viridis_c(limits = c(0,1)) + 
  facet_grid(variable~season) + 
  theme_bw() + 
  labs(x = 'Latitude', y = 'Pressure (hPa)', fill = 'R-squared', color = 'R-squared') +
  scale_y_continuous(trans = 'reverse') + 
  scale_x_continuous(breaks = c(-90, -45, 0, 45, 90)) + 
  theme(axis.text = element_text(size = 8), legend.position = 'bottom')
ggsave(paste0('src/eda/paper_plots/', params[['variable']], '_r2_lat_plev.png'),
       width = 6, height = 7.5/2)



if ('RESTOM' %in% unique(df[['variable']])) {
  df_restom <- df %>%
    filter(variable == 'RESTOM') %>%
    filter(type %in% c('surr', 'sim')) %>%
           tidyr::pivot_wider(names_from = type, values_from = value, 
                              id_cols = c('season', 'lon', 'lat', 'plev', 'sim_run', 'variable', 'res'))
  r2_restom <- summary(lm(surr~sim, data = df_restom))$r.squared
  
  ggplot(data = df_restom,
         aes(x = sim , y = surr)) + 
    geom_point(size = .5) + 
    geom_abline(intercept = 0, slope = 1) + 
    geom_smooth(se = F, method = 'lm', size = .5) + 
    theme_bw() + 
    annotate(geom = 'text', x = 2.5, y = -1, label = paste0('R-squared: ', round(r2_restom,3))) + 
    labs(x = expression('E3SM output top of atmosphere energy balance (W/m'^2 *')'),
         y = expression('Surrogate-predicted top of atmosphere energy balance (W/m'^2 *')')) + 
    theme(axis.title = element_text(size = 9))
  ggsave(paste0('src/eda/paper_plots/', params[['variable']], '_r2_global.png'),
         width = 6, height = 7.5/2)
}
```

# A look at MSE and general nature of residuals

## Center the data by the observations
```{r}
df_obs <- df  %>%
  filter(type == 'obs', res == '24x48') %>%
  rename(obs = value) %>%
  dplyr::select(-type, - sim_run)
rm(data_frame_list)
df_diff <- left_join(df, df_obs) %>%
  mutate(diff = value - obs) %>%
  left_join(area_df)
```


```{r, dpi = 288, fig.height = 16, fig.width = 15}
df_mse  <- df_diff %>%
  mutate(area_wt = ifelse(is.na(area_wt), 1, area_wt)) %>%
  group_by(season, variable, type, sim_run) %>%
  summarize(mse = mean(area_wt*diff^2, na.rm = T)) %>%
  filter(type != 'obs') %>% 
  ungroup()
 

ggplot() + 
  geom_point(data = df_mse %>% filter(!(type %in% c('sim', 'surr'))), aes(x = type,y = sqrt(mse), color =type), size = 2) + 
  geom_boxplot(data = df_mse %>% filter(type == 'sim'), aes(x = type,y = sqrt(mse)), width = .5) +
  theme(axis.text.x = element_text(angle = 90), legend.position = 'right',
        text = element_text(size = 16))+
  geom_hline(data = df_mse %>% filter(type == 'ctrl'), aes(yintercept = sqrt(mse)),
             size = .2, color = 'black') +
  labs(x = 'Type', y = 'RMSE', color = 'Type')+
  facet_grid(variable~season, scales = 'free_y')
```

# Look at the input parameters

```{r, dpi = 144, include=F, eval=F}
gc()
sim_minus_obs <- df %>%
  dplyr::filter(type == 'sim', variable != 'RESTOM') %>%
  left_join(df_obs) %>%
  mutate(diff = value - obs) %>%
  left_join(area_df)
MSE_obs <- sim_minus_obs %>%
  mutate(area_wt = ifelse(is.na(area_wt), 1, area_wt))  %>%
  group_by(sim_run, variable) %>%
  summarize(obs_error = mean(area_wt * (diff)^2,  na.rm = T))


optim_df  <- left_join(MSE_obs, data.frame(sim_run = as.character(1:250),
                                       lhs)) %>%
  tidyr::pivot_longer(cols = c('ice_sed_ai', 'clubb_c1', 'clubb_gamma_coef',
                               'zmconv_tau', 'zmconv_dmpdz'))

ggplot(data = optim_df, aes(x = value, y = obs_error))  + 
  geom_point(size = .35, stroke = .35) + 
  facet_grid(variable~name, scales = 'free') +
  labs(x = 'Parameter value', y = 'Simulation MSE compared to observations, lat/lon fields')

ggplot(data = filter(optim_df, variable == 'LWCF', name %in% c('zmconv_tau', 'zmconv_dmpdz')) %>%
         tidyr::pivot_wider(names_from = 'name', values_from = 'value'), aes(x = zmconv_tau, y = zmconv_dmpdz, color = obs_error))  + 
  geom_point() + 
  scale_color_viridis_c() 

ggplot(data = filter(optim_df, variable == 'Z500', name %in% c('zmconv_tau', 'zmconv_dmpdz')) %>%
         tidyr::pivot_wider(names_from = 'name', values_from = 'value'), aes(x = zmconv_tau, y = zmconv_dmpdz, color = obs_error))  + 
  geom_point() + 
  scale_color_viridis_c() 

par(mfrow = c(1,1))
cor_df <- optim_df %>%
  tidyr::pivot_wider(names_from = 'name', values_from = 'value') %>%
  tidyr::pivot_wider(names_from = 'variable', values_from = 'obs_error') %>%
  ungroup() %>% 
  dplyr::select(-sim_run) %>%
  cor() %>% 
  {data.frame(x = factor(rownames(.), levels = rownames(.)), y = factor(rep(colnames(.), each = nrow(.)), levels = colnames(.)), 
             value = as.vector(.))}
ggplot(data = cor_df, aes(x = x, y = y, fill = value)) +
  geom_raster() + 
  scale_fill_gradient2(limits = c(-1,1)) + 
  theme(axis.text.x = element_text(angle = 270 + 45)) +
  labs(fill  = 'Correlation')

```




